---
title: "Treatment Effects in Nested Subgroups"
author: "Jean Morrison"
date: "January 1, 2017"
output: pdf_document
bibliography: ci_bib.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.show='hold', out.width="0.5\\textwidth")
```

This document accompanies the paper "Rank conditional coverage and confidence intervals in high dimensional problems." by Jean Morrison and Noah Simon. Here we walk through the simulations in section 3.2 and reproduce the results shown in the paper. 
We make use of the `rcc` and `rccSims` packages which accompanies the paper and can be found at [github.com/jean997/rcc](https://github.com/jean997/rcc) and [github.com/jean997/rccSims](https://github.com/jean997/rccSims).

## Simulation Set-up

Here we consider data from a clinical trial examining the effect of a a treatment on an outcome $Y$. For each participant we have also collected a biomarker $W$ and suspect that the treatment has a greater effect for patients with larger values of $W$. We are interested in establishing a cut-point in $W$ to guide enrolment of a future trial so we will estimate the treatment effect in subgroups defined by the biomarker. 


The relationship between $Y$, the treatment, and the biomarker is given by
$$
E[Y | trt, W] = \begin{cases}
 0 \qquad & W < 0.5\\
\frac{1}{2}\cdot trt & W \geq 0.5
\end{cases}.
$$
For participant $i$, the observed value of $Y$ is $Y_i = E[Y | W_i, trt_i] + \epsilon$ where $\epsilon\sim N(0, 1/4)$.

In each simulation, we generate data for 400 participants --- 200 in each treatment arm. The vlaue biomarker is uniformly distributed between 0 and 1. Here we generate data for one simulation:

```{r}
library(ggplot2)
n <- 2*200
mean_outcome <- function(x, trt){
  if(x < 0.5) return(0)
  return(0.5*trt)
}
set.seed(1e7)
dat <- data.frame("trt"=rep(c(0, 1), each=n/2), "w"=runif(n=n))
dat$Ey <- apply(dat, MARGIN=1, FUN=function(z){mean_outcome(z[2],z[1])})
dat$y <- rnorm(n=n, mean=dat$Ey, sd=0.5)
dat$trt <- as.factor(dat$trt)
ggplot(dat) + geom_line(aes(x=w, y=Ey, group=trt, col=trt)) + 
  geom_point(aes(x=w, y=y, col=trt, group=trt)) + 
  ylab("E[Y]") + theme_bw() + theme(panel.grid = element_blank())
```

We will consider 100 cutopoints between 0.1 and 0.9, $w_1, \dots, w_100$. For each cutpoint, $j \in 1, \dots, 100$, we estimate the difference in average treatment effect for individuals above and below the cutpoint:
$$ \beta_j = (E[Y | trt=1, W \geq w_j] - E[Y | trt=0, W \geq w_j])- (E[Y | trt=1, W < w_j] - E[Y | trt=0, W < w_j])$$

We can estimate $\beta_j$ by fitting the parameters in the regression

$$
Y = \beta_0 + \alpha_1 trt + \alpha_2 1_{W > w_j} + \beta_j trt 1_{W > w_j} + \epsilon
$$
by least squares. Here we calculate $\hat{\beta}_j$ for each cutpoint:

```{r cutpoints}
n.cutpoints <- 100
cutpoints <- seq(0.1, 0.9, length.out=n.cutpoints)
#Indicator variables
ix <- sapply(cutpoints, FUN=function(thresh){as.numeric(dat$w >= thresh)})
#Run the linear regressions
stats <- apply(ix, MARGIN=2, FUN=function(ii){
      f <- lm(y~trt*ii, data=dat)
      summary(f)$coefficients[4, 1:3]
    })
stats <- data.frame(matrix(unlist(stats), byrow=TRUE, ncol=3))
names(stats) <- c("beta", "se", "tstat")
stats$cutpoint <- cutpoints
j <- order(abs(stats$tstat), decreasing=TRUE)
stats$rank <- match(1:nrow(stats), j)
head(stats)
```

The true value of $\beta_j$ is 
$$0.5*\frac{min(1-w_j, 0.5)}{(1-w_j)} - 0.5\frac{max(0, w_j-0.5)}{w_j}$$
```{r truth}
stats$truth <-sapply(cutpoints, FUN=function(wj){
  0.5*(min(1-wj, 0.5)/(1-wj)) - 0.5*max(0, wj-0.5)/wj})
```

Plotting effect size estimates and true parameter values vs. cutpoint:
```{r echo=FALSE}
library(tidyr)
stats_long <- gather(stats[, c("cutpoint", "beta", "truth")], "type", "value", -cutpoint)
ggplot(stats_long) + geom_point(aes(x=cutpoint, y=value, group=type, color=type)) + 
  scale_color_discrete(labels=c("Estimate", "Truth")) + ggtitle("Effect Size Estimates") + 
  theme_bw() + theme(panel.grid=element_blank(), axis.title.y=element_blank(), legend.title = element_blank())
```

Because we use the same individuals to estimate each parameter, the estimates are highly correlated. 

## Confidence Intervals 

### Non-Parametric Bootstrap
The non-parametric bootstrap is the most appropriate choice for these data since the estimates are correlated. Only the non-parametric bootstrap can model this correlation. To use the `nonpar_bs_ci` function in the `rcc` package to compute the nonparametric boostrap confidence intervals, we must supply an analysis function. We will use a data object that has $trt$, $w$, $E[Y]$, and $Y$ as the first four columns and the indicator variables $1_{W > w_j}$ as the next 100 columns as input.

```{r, af, cache=TRUE}
mydata <- cbind(dat, ix)
analysis.func <- function(data){
  y <- data[,4]
  trt <- data[,1]
  stats <- apply(data[, 5:(n.cutpoints + 4)], MARGIN=2, FUN=function(ii){
      f <- lm(y~trt*ii)
      if(nrow(summary(f)$coefficients) < 4) return(rep(0, 3))
      summary(f)$coefficients[4, 1:3]
    })
  stats <- data.frame(matrix(unlist(stats), byrow=TRUE, ncol=3))
  names(stats) <- c("estimate", "se", "statistic")
  return(stats)
}
```


Here we calculate the non-parametrci bootstrap confidence intervals (Algorithm 3 in the paper):
```{r, nonpar, cache=TRUE}
library(parallel)
ci.nonpar <- nonpar_bs_ci(data=mydata, analysis.func=analysis.func, n.rep=500, parallel=TRUE)
sum(ci.nonpar$ci[,1] <= stats$truth & stats$truth <= ci.nonpar$ci[,2])/n.cutpoints
```


Here are the intervals plotted versus rank. Colored points indicate the true value of the parameter.
```{r}
plot_cis(rank, ci.nonpar$ci, stats$truth, plot.truth=TRUE) + 
  ggtitle("Non-Parametric Bootstrap Confidence Intervals")
```
Here they are plotted versus cutpoint:
```{r}
plot_cis(cutpoints, ci.nonpar$ci, stats$truth, plot.truth=TRUE) + 
  ggtitle("Non-Parametric Bootstrap Confidence Intervals")
```




### Parametric Bootstrap

We could instead use the parametric bootstrap (Algorithm 2 and Supplemental Algorithm 2):
```{r parboot}
ci.par <- rcc::par_bs_ci(beta=stats$beta, se=stats$se, n=500, use.abs=TRUE)
```
Here are the intervals plotted versus rank. Colored points indicate the true value of the parameter.
```{r}
plot_cis(rank, ci.par$ci, stats$truth, plot.truth=TRUE) + 
  ggtitle("Parametric Bootstrap Confidence Intervals")
```
Here they are plotted versus cut-point:
```{r}
plot_cis(cutpoints, ci.par$ci, stats$truth, plot.truth=TRUE) + 
  ggtitle("Parametric Bootstrap Confidence Intervals")
```


### Marginal Confidence Intervals
For comparison, let's look at the naive confidence intervals:
```{r}
 ci.naive <- cbind(stats$beta-stats$se*qnorm(0.95), stats$beta + stats$se*qnorm(0.95))
sum(ci.naive[,1] <= stats$truth & stats$truth <= ci.naive[,2])/100
```
```{r}
plot_cis(rank, ci.naive, stats$truth, plot.truth=TRUE) + 
  ggtitle("Standard Marginal Confidence Intervals")
plot_cis(cutpoints, ci.naive, stats$truth, plot.truth=TRUE) + 
  ggtitle("Standard Marginal Confidence Intervals")
```
These do pretty well since the estimates are so correlated.
## Selection Adjusted Confidence Intervals

## Empirical Bayes Confidence Intervals

## Replicate the results in the paper

```{r, eval=FALSE}
library(parallel)
biomarker_results <- biomarker_sim(n=400, n.rep = 100, n.cutpoints = 100, seed =1e7)
```


Plotting:
```{r plotresults, show="hold", out.width="0.5\\textwidth"}
data("biomarker_results", package="rcc")
coverageplot <- rcc:::plot_coverage(biomarker_results, proportion=1,
      cols=c("black",  "deeppink3",  "red", "gold4", "forestgreen", "purple"),
      simnames=c("naive",  "par",    "nonpar", "ash", "wfb", "selInf1"), 
      ltys= c(2, 1, 3, 6, 4, 2), 
      span=0.5, main="Rank Conditional Coverage", y.range=c(-0.02, 1.02),
      legend.position = "none") + theme(plot.title=element_text(hjust=0.5))
widthplot <- rcc:::plot_width(biomarker_results, proportion=1,
      cols=c("black",  "deeppink3",  "red", "gold4", "forestgreen", "purple"),
      simnames=c("naive",  "par",    "nonpar", "ash", "wfb", "selInf1"), 
      ltys= c(2, 1, 3, 6, 4, 2), span=0.5, main="Interval Width",
      legend.position = "none") + theme(plot.title=element_text(hjust=0.5))
legend <- rcc:::make_sim_legend(legend.names = c("Marginal", "Parametric\nBootstrap", 
                                                 "Non-Parametric\nBootstrap", "ashr", "WFB", "RTT"), 
            cols=c("black",  "deeppink3",  "red", "gold4", "forestgreen", "purple"),
            ltys= c(2, 1, 3, 6, 4, 2))
coverageplot
widthplot
legend$plot
```

```{r, echo=FALSE, eval=FALSE}
h1 <- grid.arrange(coverageplot, widthplot, ncol=2)
ggsave(h1, file="~/Dropbox/rcc-temp/for_jcgs/img/biomarker.pdf", heigh=4, width=8, units="in", dpi=300)
```
